---
title: Process Mining
summary: On using data analysis to understand business processes
tags:
- Process Mining
date: "2020-08-10T00:00:00Z"

# Optional external URL for project (replaces project detail page).
external_link: ""

image:
  caption: Photo by rawpixel on Unsplash
  focal_point: Smart


url_code: ""
url_pdf: ""
url_slides: ""
url_video: ""

# Slides (optional).
#   Associate this project with Markdown slides.
#   Simply enter your slide deck's filename without extension.
#   E.g. `slides = "example-slides"` references `content/slides/example-slides.md`.
#   Otherwise, set `slides = ""`.
slides: example
---

Big Processes are a Big Topic in data analytics. This is particularly true in enterprises who are placing a heavy emphasis on operational efficiency and are looking to move towards a culture of continual improvement.

One of the ways businesses are tackling process analysis is via techniques like "process mining". 

Process mining is a a set of analytical techniques we can use to gain insights into the following types of questions:

- Process discovery - what is really happening ? How can we understand the current process, especially where no formal model artefacts may exist ?

- How often are activities performed ?

- Who are the fastest at performing certain tasks ?

- What is the "cycle" time ? What is the "waiting" time ?

- How often do "handoffs" occur ?

- ... and many other questions !

Tools like R provide an ideal platform for getting into a journey of process mining.

There is a brilliant R library centered on process mining from the people here:

  Gert Janssenswillen (2020). bupaR: Business Process Analysis in R. R package
  version 0.4.4. https://CRAN.R-project.org/package=bupaR

I'm going to use bupaR and its suite of related packages to explore a dataset of transactional event log type data to learn a bit about the processes associated this data. I've synthesised this data for demonstration purposes. 

Business process management (BPM) or typical workflow type applications which are ubiquitous in corporate settings are a common candidate for this type of data supply for analysis.

First up, get the data from an Excel file I created earlier:

```{r get_the_data, warning=FALSE, message=FALSE}

library(readxl)

auditextract <- read_excel("G:/OneDrive/Documents/bupaR/AuditExtract1.xlsx")

```

Then, load up the required bupaR libraries and create an event log object from the source while mapping the relevant columns into the event log construct in bupaR:

```{r create_event_log, warning=FALSE, message=FALSE}

library(bupaR)
library(processanimateR)

el <- auditextract %>%
  mutate(status = "complete",
         activity_instance = 1:nrow(.)) %>%
  eventlog(
    case_id = "AuditCaseID",
    activity_id = "StatusDescription",
    activity_instance_id = "rowid",
    lifecycle_id = "status",
    timestamp = "CreatedTimestamp",
    resource_id = "ResourceName2"
  )

# animate_process(eldata)
# nrow(el)

```

So, what does the raw process consist of ? With bupaR, we can easily visualise a process by plotting a process map. Let's have a look:

```{r}
process_map(el)
```

Well, stop right there ! The obvious comes to mind straight away - this ain't no simple map. If we want to break it down and get some insights into what's really going on in the workflow, let's ...

More text 

