---
title: Process Mining - Part 2
summary: Part 2 on the topic of Process Mining
tags:
- Process Mining
date: "2020-08-10T00:00:00Z"

# Optional external URL for project (replaces project detail page).
external_link: ""

image:
  caption: Photo by rawpixel on Unsplash
  focal_point: Smart


url_code: ""
url_pdf: ""
url_slides: ""
url_video: ""

# Slides (optional).
#   Associate this project with Markdown slides.
#   Simply enter your slide deck's filename without extension.
#   E.g. `slides = "example-slides"` references `content/slides/example-slides.md`.
#   Otherwise, set `slides = ""`.
slides: example
---
This is a continuation from [Part 1](/project/internal-project-2/) on the topic of process mining. 

```{r get_the_data, echo=FALSE, warning=FALSE, message=FALSE}

library(readxl)
auditextract <- read_excel("G:/OneDrive/Documents/bupaR/AuditExtract1.xlsx")

library(bupaR)
library(processanimateR)
library(edeaR)

el <- auditextract %>%
  mutate(status = "complete",
         activity_instance = 1:nrow(.)) %>%
  eventlog(
    case_id = "AuditCaseID",
    activity_id = "StatusDescription",
    activity_instance_id = "rowid",
    lifecycle_id = "status",
    timestamp = "CreatedTimestamp",
    resource_id = "ResourceName2"
  )

library(dplyr)
qs <- auditextract %>% distinct(objectTypeDesc, StatusDescription)

elc <- el %>%
  act_collapse(StartApplication = filter(qs, objectTypeDesc=='Start Application')$StatusDescription) %>%
  act_collapse(UploadDocuments = filter(qs, objectTypeDesc=='Upload Documents')$StatusDescription) %>%
  act_collapse(Settlement = filter(qs, objectTypeDesc=='Settlement')$StatusDescription) %>%
  act_collapse(ProcessDocuments = filter(qs, objectTypeDesc=='Process Documents')$StatusDescription) %>%
  act_collapse(Disburse = filter(qs, objectTypeDesc=='Disburse')$StatusDescription) %>%
  act_collapse(StartDocuments = filter(qs, objectTypeDesc=='Start Documents')$StatusDescription) 

```

One of bupaR's coolest features is the ability to **animate** a process. The animation represents the movement of cases through the processes, based on the timing data. Check this out:

```{r animate_1, warning=FALSE, message=FALSE}
animate_process(elc)
```

In the following sections, I'll list a quick rundown of a few more features from bupaR. Really, this is just scratching the surface of what bupaR and process mining can provide - I encourage you if you're interested in this topic to check out bupaR and start a journey of your own.

We can add a measure of **performance** in terms of timing to the process map. Here I've added median minutes for the activities, with a view of relative frequencies (replacing the absolute frequencies we saw earlier). We can see clearly now that the movement from "Upload Documents" to "Process Documents" is the heaviest consumer of the timing at over 78 minutes:

```{r process_map_performance, warning=FALSE, message=FALSE}
elc %>%
  process_map(type_nodes = frequency("relative_case"), performance(median, "mins", color_scale = "Purples"))

```

How is the **workload** distributed amongst the team ? Let's have a look at just that Settlement activity and see:

```{r process_map_resource1, warning=FALSE, message=FALSE}
el %>%
  filter(objectTypeDesc == 'Settlement') %>%
  resource_involvement("resource") %>%
  plot()
```
 
In bupaR, "traces" show the different activity sequences that occur. It may be of interest to view both the most common and least common sequences. Here we will narrow down again to just that Settlement activity to see how the statuses move in different sequences within there:

```{r process_map_traces1, warning=FALSE, message=FALSE}
el %>%
  filter(objectTypeDesc == 'Settlement') %>%
  trace_explorer(coverage = 0.5)
```


Thank you for reading !

